Sketch Recognition and Understanding

### 2009-09-26 draft/System.tex

The contribution in the system is mainly in attempting to generate the optimal segmentation using PSO which exhibited superiority with respect to similar algorithms in various other applications \cite{PolygonApproximationPSO}. The system attempts to find the optimal decomposition of the input stroke using PSO with the help of curvature and speed information. The use of PSO algorithm helps in eliminating the effects of input noise, problems of over and under segmentation that were in pervious systems. The proposed symbol recognizer uses the generated segmentation to compute the feature vector which is used to identify more complex symbols. This enables the system to be scalable for more complex shapes.

@@

Also the time difference between samples was used in \cite{polygonfeedback31} as it provides more distinct maxima than the speed information. Agar et al. \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena. The points are sampled at regular time intervals while the pointing device is moving. There are no points while the pointing device is stationary. This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary. Contrary to speed information, where the users draw with variant speeds the speed information has a lot of noise in the data. Figure \ref{fig:speed2Distance} shows the time difference and the speed graphs for the stroke drawn in Figure \ref{fig:orignalStroke}.  Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}. Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}

@@

The system uses a composite set of features which include global shape properties like Rubine feature set \cite{gestureexample12},  ink density \cite{GeometryAndDomain102}, some appearance based properties like Zernike moments \cite{HeloiseBeautification}, and some stroke based structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol. The following list gives details on all the features used in the system.

\subsection{Structural and geometrical Features(FS1)}
Features defines the structure of the geometrical symbol. Section \ref{sec:steprec} descripe in details how these feature are computed. 
%\begin{description}
% \item [Structural and geometrical Features (FS1)] 
  \begin{itemize}
	 \item \emph{Segments:} Number of segments in the symbol.
	 \item \emph{Strokes:} Number of strokes or partial strokes that created the symbols.  
		\item  \emph{Primitives:} Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item \emph{Curves:} Number of curves or ellipses in the symbol. 
		\item \emph{Lines:} Number of lines in the symbol. 
		\item \emph{Perpendicular} lines Number of perpendicular lines. 
		\item \emph{Parallel} lines Number of parallel lines. 
		\item \emph{Intersections} Number of intersection between lines and curves. 
		\item \emph{T intersections} Number of T intersections. 
		\item \emph{L intersections} Number of L intersections. 
		\item \emph{X intersections} Number of X intersections.

		%\item[No. of holes] Number of holes in the symbols. 
\end{itemize}
%\begin{description}
%	 \item[Segments] Number of segments in the symbol.
%	 \item [Strokes] Number of strokes or partial strokes that created the symbols.  
%		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
%		\item [Curves] Number of curves or ellipses in the symbol. 
%		\item [Lines] Number of lines in the symbol. 
%		\item [Perpendicular lines] Number of perpendicular lines. 
%		\item [Parallel lines] Number of parallel lines. 
%		\item [Intersections] Number of intersection between lines and curves. 
%		\item [T intersections] Number of T intersections. 
%		\item [L intersections] Number of L intersections. 
%		\item [X intersections] Number of X intersections.
%		
%		
%		%\item[No. of holes] Number of holes in the symbols. 
%
%\end{description}
\subsection{Rubine Feature Set (FS2)} Features introduced by Rubine\cite{gestureexample12} for single stroke gestures. These features represent the global shape of the symbol drawn. To compute the features we append all segments points into single path $ink_{path}$. The features is computed on the generated path $ink_{path}$.%then compute the features based on this path.  
\begin{enumerate}
	\item Cosine of starting angle.
	\item Sine of starting angle.
	\item Length of diagonal of bounding box. It gives an idea of the size of the bounding box).
	\item Angle of diagonal. It gives an idea of the shape of the bounding box (long, tall, square).
	\item Distance from start to end.  
	\item Total stroke length
	\item Change in Rotation( Arctan ). It gives the directional angle.
	\item Absolute rotation 
	\item Rotation squared 
	\item The maximum speed reached (squared) 
	\item Total time of stroke 
\end{enumerate}
\subsection{Statistical Features (FS3)} 
\begin{enumerate}
	\item \textsl{Zernike moments:} Zernike moments of order n\footnote{ We tried different values of n. Results show the analysis of each order.} \cite{HeloiseBeautification}. 
\end{enumerate}
\subsection{Global shape properties set (FS4)}
 Features that are composites of different geometrical and statistical symbol. These features represent the global shape of the symbol drawn.

 	\begin{itemize}
\item \emph{Size Ratio} Ratio between width to height of the symbol.
	\item \emph{Ink density} Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
 	\item \emph{Convex Hull Area} Area of convex hull with respect to area of bounding box of symbol.
	\item \emph{Convex Hull Perimeter} Perimeter of convex hull with respect to total length of symbol.
		\item \emph{Mean Centroidal radius} The Mean of the centroidal radius which is the distance from each point in the symbol to the center of gravity.

	\item \emph{Mean Time difference} The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
	%\item [Ra]
  \end{itemize}

@@

\subsection{Support Vector Machines}
\label{Sec:SVMdetail}
The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{svmintroduce} in 1995. It is gaining increase popularity due to the high performance and efficient implementation. The goal of SVM algorithms is to generate a separating plane between two different classes which are represented using a set of n-dimensional vectors. The generated plane is used as a good classifier to the unseen examples. However, there can be more than one plane that can separates the data therefore SVM attempts to find the plane that maximize the margin (distance between classifier and nearest sample) between the two classes \ref{fig:svm1}. To calculate the margin, two different hyper planes are constructed, one plane on each side of the data, these two planes are then moved to generate the maximum margin possible.  
\begin{figure}
	\centering
		\includegraphics{afterDefense/svm1.jpg.eps}
	\caption{Different Classifier planes }
	\label{fig:svm1}
\end{figure}

@@

The main problem of SVM is constructing the plane that is furthest from both classes. The main approach is to use two parallel supporting planes and maximize the margin between them. All point in a class should be in one side of the plane in order to be a supporting plane.  Therefore, for the points with the class label $+1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i\> b$ or $\omega \cdot x_i -b \> 0$%, and for the points with class label $-1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i \< b $.
Suppose that smallest value of $\mid \omega \cdot x_i -b\mid$ is $\kappa$, then when $ \omega \cdot x_i -b \geq \kappa$. The argument inside the decision function is invariant under positive rescaling so we will implicitly fix a scale by requiring $\omega \cdot x_i -b \geq 1 $. Similarly the points with class label $-1$ requires $\omega \cdot x_i -b \leq -1 $. To find the plane furthest from both classes, we need to maximize the distance between the support planes for each class as illustrated in Figure \ref{fig:svm2}. To accomplish this, both planes are pushed apart until they bump into small number of data points ( the support vectors ) for each class. Figure \ref{fig:svm2} shows the support vectors as filled data points, 
\begin{figure}
	\centering
		\includegraphics{afterDefense/svm2.jpg.eps}
	\caption{Margin of Support planes }
	\label{fig:svm2}
\end{figure}

@@

\begin{figure}
	\centering
		\includegraphics{afterDefense/svm5.jpg.eps}
	\caption{Linearly Inseparable}
	\label{fig:svm5}
\end{figure}
A minor change in the supporting plane QP problem can accomplish this. A non negative slack or error variable $z_i$ is added to each constraint.  A weighted penalty term is also added to the objective as follows:
 \begin{equation}
 \begin{array}{c}
\min_{w,b,z}  \quad \frac{{\parallel \omega \parallel}^2}{2}+C \sum_{i=1}^{l}z_i  \\
s.t.  \quad y_i(\omega \cdot x_i -b)+z_i \geq 1  \\
    z_i \geq 0  i=1,\cdots,m
\end{array} 
\label{eq:inseperable}
\end{equation}

@@

\begin{figure}
	\centering
		\includegraphics{afterDefense/svm6.jpg.eps}
	\caption{non Lineary problem }
	\label{fig:svm6}
\end{figure}

@@

\begin{equation}
\begin{array}{c}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_j\vartheta(x_i)\vartheta(x_j)}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
 C \geq \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
 \end{array} 
\label{eq:dualkernel}
\end{equation}
Notice that the mapped data only occurs as inner product in the objective. Now we apply a little mathematically magic known as hilbern schmidt kernels, first applied to SVMs in \cite{svmintroduce}.  By Mercer's theorem, we know that for certain mapping $\theta$ and any two points $u$ and $v$, the inner product of the mapped points can be evaluated using the kernel function without ever explicitly knowing the mapping, e.g., $\theta(u)\theta(v)\equiv K(u,v)$. Some of the more popular known kernels are given below. New kernels are being developed to fit domain specific requirements. 
\begin{center}
	\begin{tabular}{cc}
	$\vartheta(u)$  & $K(u,v)$ \\ \hline 
	Degree $d$ polynomial & $\left(u\cdot v +1\right)^d$ \\ 
	Radial Basis Function & $\exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) $ \\ 
	Two layer Neural Network& $ sigmoid \left(\eta\left(u\cdot v +1\right)+c\right)$ \\ 
	\end{tabular}
\end{center}
To change from a linear to nonlinear classifier, one must only substitute a kernel evaluation in the objective instead of the original dot product. Thus by changing kernels we can get different highly nonlinear classifiers. No algorithmic changes are required from linear case other than substitute of a kernel evaluation for simple dot product. All the benefits of the original linear SVM methods are maintained. We can train a highly nonlinear classification function such as a polynomial or a radial basis function machine, or a sigmoid neural network using robust, efficient algorithms that have no problem with local minima. By using kernel substitution a linear algorithm (only capable of handling separable data) can be turned into a general nonlinear algorithm. 
%\begin{equation}
%\begin{array}{cc}
%\underline{&}\\  
% & \left(u\cdot v +1\right)^d \\
%Radial Basis Function & \exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) \\
%Two layer Neural Network & sigmoid \left(\eta\left(u\cdot v +1\right)+c\right) 
% \end{array} 
%\label{eq:diffkernel}
%\end{equation}

%The foundations of Support Vector Machines (SVM) have
\subsubsection{SVM  Classification}
The resulting SVM method can be summarized as follows 
\begin{enumerate}
	\item Selecting the SVM classifier Paramters. Parameter $C$ representing the tradeoff between minimizing the training set error and maximizing the margin. 
	\item Select A kernel function and their kernel parameters. For example for the radial basis function kernel, one must select width of the Gaussian $\gamma$. 
	\item Solve dual QP Equation \ref{eq:dualkernel} or an alternative SVM formulation using an appropriate quadratic programming or linear programming algorithm. 
	\item Get the support vector and recover the emprial threshold  variable $b$. 
	\item Classify new points using $f(x)=sign\left(\sum{i}y_i\alpha_i K\left(x,x_i\right)-b\right)$
\end{enumerate}
The parameters in step 1. are selected using cross validation if sufficient data are available. %However, recent model  selectin strategies can 
\subsubsection {Multi Class Classifiers}

 There is mainly two popular methods to change the binary SVM classification problem into Multiclass classification problem. The first methods is One Versus All (OVA) Strategy where M binary classifiers is constructed one classifier for each class.  The $ith$ classifier is trained using positive example form $ith$ class and negative example from all other classes. The output function $p_i$ represents the distance and classification of the example from the $ith$ classifier. For a new example $x$, OVA SVM strategy assigns it to the class with the largest value of $p_i$.

@@

In this step the user draw one or more stroke to draw a symbol. Figure \ref{fig:clock} illustrate the sequence and direction of user draw. In this section, each stroke user draws will be labeled as it illustrated in Figure \ref{fig:clockstroke1} and Figure \ref{fig:clockstroke2}.  Table \ref{tab:StepsStroke1} shows the detailed output of the program at each step. The input of this step is the a set of strokes, each stroke has a set of  points and the output is a list of segments that will be used in the recognition process. More detailed example and algorithms are demonstrated in the Appendex \ref{ChapterstepExample} \\

@@

\begin{enumerate}
\item Extracting $P_dp$ as in Section \ref{sec:Preprocessing} and Algorithm \ref{extractpdp}.
\item Ellipse Fitting: which is an attempt to fit as an ellipse as in details in Section \ref{sec:EllipseDetection} and Algorithm \ref{ellipseFitAlg} in Appendex \ref{ChapterstepExample}.
\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}. The alogrithm will be similar to the PSO algorithm  in section \ref{sec:ParticleSwarmAlgorithm}. (see Algorithm \ref{AlgS1alg} in Appendex \ref{ChapterstepExample})
%he axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
\item Segmentation ALS2 as explained in Section \ref{sec:PolygonDivisionAlgorithm}. ( see Algorithm \ref{AlgS2alg} in Appendex \ref{ChapterstepExample} ). 
\item Append best segmentation to List of segmentation ($L_s$) of this symbols. 
\end{enumerate}

\begin{landscape}
The steps of the first stroke (details of the ouputs are in  Appendex \ref{ChapterstepExample} in Table \ref{tab:StepsStroke1Details}):
\begin{scriptsize}
%\begin{table}
	%\centering 
%\scalebox{0.6}{		
\begin{longtable}{|p{2cm}|p{2cm}|p{5cm}|p{2cm}|p{5cm}|}
\caption{ Output of System in Each Step of Segmentation of First Stroke}
\label{tab:StepsStroke1} \\

@@

 \includegraphics[scale=0.5]{afterDefense/ellipse.jpg.eps}
 \\ \hline
 	\end{longtable}
 	\newpage
 	The steps of the second stroke:(details of the ouputs are in  Appendex \ref{ChapterstepExample} in Table \ref{tab:StepsStroke1Details})

@@

\begin{enumerate}
	\item Computing Structural and geometrical Features (FS1). This step computes number of parallel, perpendicular line and similar structural features [10 features]. Algorithm \ref{FeatureAlg} in Appendex \ref{ChapterstepExample}  is used to determine these features.
 	\item Rubine Feature Set (FS2) [12 features] .  Compute Rubine features as in \cite{gestureexample12} 
	\item Statistical Features (FS3) compute the moments as in \cite{zernike61} for N=10 there are 32 features.
	\item Global shape properties set (FS4). Firstly compute the convex hull of the stroke points then compute the rest of the features [12 features]. 
	\item Introduce the computed feature vector to the classifier and check the label of the detected symbol. 
\end{enumerate}

@@

More details are in Table \ref{tab:StepsRegDetails} in Appendex \ref{ChapterstepExample}

>>>>

The contribution in the system is mainly in attempting to generate the optimal segmentation using PSO which exhibited superiority with respect to similar algorithms in various other applications \cite{PolygonApproximationPSO}. The system attempts to find the optimal decomposition of the input stroke using PSO with the help of curvature and speed information. The use of PSO algorithm helps in eliminating the effects of input noise, problems of over and under segmentation that were in previous systems. The proposed symbol recognizer uses the generated segmentation to compute the feature vector which is used to identify more complex symbols. This enables the system to be scalable for more complex shapes.

@@

Also the time difference between samples was used in \cite{polygonfeedback31} as it provides more distinct maximum than the speed information. Agar et al. \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena. The points are sampled at regular time intervals while the pointing device is moving. There are no points while the pointing device is stationary. This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary. Contrary to speed information, where the users draw with variant speeds the speed information has a lot of noise in the data. Figure \ref{fig:speed2Distance} shows the time difference and the speed graphs for the stroke drawn in Figure \ref{fig:orignalStroke}.  Similarly, direction information is used as it provides better distinctive maximum for the corners than the curvature information \cite{meanshift10}. Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}

@@

The system uses a composite set of features (total of 70 features) which include global shape properties like Rubine feature set \cite{gestureexample12},  ink density \cite{GeometryAndDomain102}, some appearance based properties like Zernike moments \cite{HeloiseBeautification}, and some stroke based structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol. The following list gives details on all the features used in the system.

\subsection{Structural and geometrical Features(FS1)}
 Features defines the structure of the geometrical symbol. Section \ref{sec:steprec} describe in details how these 13 feature are computed.  
%\begin{description}
% \item [Structural and geometrical Features (FS1)] 
  \begin{itemize}
	 \item \emph{Segments:} Number of segments in the symbol.
	 \item \emph{Strokes:} Number of strokes or partial strokes that created the symbols.  
		\item  \emph{Primitives:} Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item \emph{Curves:} Number of curves or ellipses in the symbol. 
		\item \emph{Lines:} Number of lines in the symbol. 
		\item \emph{Perpendicular} lines Number of perpendicular lines. 
		\item \emph{Parallel} lines Number of parallel lines. 
		\item \emph{Intersections} Number of intersection between lines and curves. 
		\item \emph{T intersections} Number of T intersections. 
		\item \emph{L intersections} Number of L intersections. 
		\item \emph{X intersections} Number of X intersections.
		\item \emph{Min Radius} Minimum Radius of all curves in the symbol.
		\item \emph{Max Radius} Maximum Radius of all curves in the symbol.
		%\item[No. of holes] Number of holes in the symbols. 
\end{itemize}
%\begin{description}
%	 \item[Segments] Number of segments in the symbol.
%	 \item [Strokes] Number of strokes or partial strokes that created the symbols.  
%		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
%		\item [Curves] Number of curves or ellipses in the symbol. 
%		\item [Lines] Number of lines in the symbol. 
%		\item [Perpendicular lines] Number of perpendicular lines. 
%		\item [Parallel lines] Number of parallel lines. 
%		\item [Intersections] Number of intersection between lines and curves. 
%		\item [T intersections] Number of T intersections. 
%		\item [L intersections] Number of L intersections. 
%		\item [X intersections] Number of X intersections.
%		
%		
%		%\item[No. of holes] Number of holes in the symbols. 
%
%\end{description}
\subsection{Rubine Feature Set (FS2)} Features introduced by Rubine\cite{gestureexample12} for single stroke gestures. These 12 features represent the global shape of the symbol drawn. To compute the features we append all segments points into single path $ink_{path}$. The features is computed on the generated path $ink_{path}$.%then compute the features based on this path.  
\begin{enumerate}
	\item Cosine of starting angle.
	\item Sine of starting angle.
	\item Length of diagonal of bounding box. It gives an idea of the size of the bounding box).
	\item Angle of diagonal. It gives an idea of the shape of the bounding box (long, tall, square).
	\item Distance from start to end.  
	\item Total stroke length
	\item Change in Rotation( Arctan ). It gives the directional angle.
	\item Absolute rotation 
	\item Rotation squared 
	\item The maximum speed reached (squared) 
	\item Total time of stroke 
\end{enumerate}
\subsection{Statistical Features (FS3)} 
\begin{enumerate}
	\item \textsl{Zernike moments:} Zernike moments of order $n$\footnote{ We tried different values of n. Results show the analysis of each order.} \cite{HeloiseBeautification}.  For $n=10$ there are 32 moment features. 
\end{enumerate}
\subsection{Global shape properties set (FS4)}
 Features that are composites of different geometrical and statistical symbol. These 13 features represent the global shape of the symbol drawn.

 	\begin{itemize}
\item \emph{Size Ratio} Ratio between width to height of the symbol.
		\item \emph{Log of size Ratio} Log of the ratio between width and height of symbol.
	\item \emph{Ink density} Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
 	\item \emph{Convex Hull Area} Area of convex hull with respect to area of bounding box of symbol.

	\item \emph{Convex Hull Perimeter} Perimeter of convex hull with respect to total length of symbol.
		\item \emph{Convex Hull Point} Number of points of convex hull with respect to Number of points of symbol.
		\item \emph{Mean Centroidal radius} The Mean of the centroidal radius which is the distance from each point in the symbol to the center of gravity.
				\item \emph{Center of Gravity} The center of gravity of the symbol (both in y and x direction).
					\item \emph{Mean Time} The mean of the time values between each two successive points in the symbol. %Different strokes are appended to construct a single path.
	%\item [Ra]

	\item \emph{Mean Time difference} The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
	%\item [Ra]		
			\item \emph{Log of Area} Log of area of the symbol.
	\item \emph{Diff Area} Absolute difference of the area of bounding box and area of symbol.

  \end{itemize}

@@

\subsection{Support Vector Machines}
\label{Sec:SVMdetail}
The foundation of Support Vector Machines (SVM) were developed by Vapnik \cite{svmintroduce} in 1995. It is gaining increase popularity due to the high performance and efficient implementation. The goal of SVM algorithms is to generate a separating plane between two different classes which are represented using a set of n-dimensional vectors. The generated plane is used as a good classifier to the unseen examples. However, there can be more than one plane that can separates the data therefore SVM attempts to find the plane that maximize the margin (distance between classifier and nearest sample) between the two classes \ref{fig:svm1}. To calculate the margin, two different hyper planes are constructed, one plane on each side of the data, these two planes are then moved to generate the maximum margin possible.  
\begin{figure}
	\centering
		\includegraphics{afterDefense/svm1.jpg.eps}
	\caption{Different Classifier Planes }
	\label{fig:svm1}
\end{figure}

@@

The main problem of SVM is constructing the plane that is furthest from both classes. The main approach is to use two parallel supporting planes and maximize the margin between them. All point in a class should be in one side of the plane in order to be a supporting plane.  Therefore, for the points with the class label $+1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i\> b$ or $\omega \cdot x_i -b \> 0$%, and for the points with class label $-1$ we would like there exist $\omega$ and $b$ such that $\omega \cdot x_i \< b $.
Suppose that smallest value of $\mid \omega \cdot x_i -b\mid$ is $\kappa$, then when $ \omega \cdot x_i -b \geq \kappa$. The argument inside the decision function is invariant under positive rescaling so we will implicitly fix a scale by requiring $\omega \cdot x_i -b \geq 1 $. Similarly the points with class label $-1$ requires $\omega \cdot x_i -b \leq -1 $. To find the plane furthest from both classes, we need to maximize the distance between the support planes for each class as illustrated in Figure \ref{fig:svm2}. To accomplish this, both planes are pushed apart until they bump into small number of data points ( the support vectors ) for each class. Figure \ref{fig:svm2} shows the support vectors as filled data points, 
\begin{figure}
	\centering
		\includegraphics{afterDefense/svm2.jpg.eps}
	\caption{Margin of SVM Support planes }
	\label{fig:svm2}
\end{figure}

@@

\begin{figure}
	\centering
		\includegraphics{afterDefense/svm5.jpg.eps}
	\caption{Example of SVM Linearly Inseparable Problem}
	\label{fig:svm5}
\end{figure}
A minor change in the supporting plane QP problem can accomplish this. A non negative slack or error variable $z_i$ is added to each constraint.  A weighted penalty term is also added to the objective as follows:
 \begin{equation}
 \begin{array}{c}
\min_{w,b,z}  \quad \frac{{\parallel \omega \parallel}^2}{2}+C \sum_{i=1}^{l}z_i  \\
s.t.  \quad y_i(\omega \cdot x_i -b)+z_i \geq 1  \\
    z_i \geq 0  i=1,\cdots,m
\end{array} 
\label{eq:inseperable}
\end{equation}

@@

\begin{figure}
	\centering
		\includegraphics{afterDefense/svm6.jpg.eps}
	\caption{Example of an non Linearly SVM Problem }
	\label{fig:svm6}
\end{figure}

@@

\begin{equation}
\begin{array}{c}
\min_{\alpha} \ \ \   \frac{1}{2} \sum_{i=1}^{m}{\sum_{j=1}^{m}{y_iy_j\alpha_i\alpha_j\vartheta(x_i)\vartheta(x_j)}} - \sum_{i=1}^{m}{\alpha_i} \\
s.t. \ \ \sum_{i=1}^{m}{y_i\alpha_i}=0   \\
 C \geq \alpha_i \geq 0  \ \ \ i=1,\cdots , m 
 \end{array} 
\label{eq:dualkernel}
\end{equation}
Notice that the mapped data only occurs as inner product in the objective. Now we apply a little mathematically magic known as Hilbert Schmidt kernels, first applied to SVM in \cite{svmintroduce}.  By Mercer's theorem, we know that for certain mapping $\theta$ and any two points $u$ and $v$, the inner product of the mapped points can be evaluated using the kernel function without ever explicitly knowing the mapping, e.g., $\theta(u)\theta(v)\equiv K(u,v)$. Some of the more popular known kernels are given below. New kernels are being developed to fit domain specific requirements. 
\begin{center}
	\begin{tabular}{cc}
	$\vartheta(u)$  & $K(u,v)$ \\ \hline 
	Degree $d$ polynomial & $\left(u\cdot v +1\right)^d$ \\ 
	Radial Basis Function & $\exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) $ \\ 
	Two layer Neural Network& $ sigmoid \left(\eta\left(u\cdot v +1\right)+c\right)$ \\ 
	\end{tabular}
\end{center}
To change from a linear to nonlinear classifier, one must only substitute a kernel evaluation in the objective instead of the original dot product. Thus by changing kernels we can get different highly nonlinear classifiers. No algorithmic changes are required from linear case other than substitute of a kernel evaluation for simple dot product. All the benefits of the original linear SVM methods are maintained. We can train a highly nonlinear classification function such as a polynomial or a radial basis function machine, or a sigmoid neural network using robust, efficient algorithms that have no problem with local minima. By using kernel substitution a linear algorithm (only capable of handling separable data) can be turned into a general nonlinear algorithm. 
%\begin{equation}
%\begin{array}{cc}
%\underline{&}\\  
% & \left(u\cdot v +1\right)^d \\
%Radial Basis Function & \exp \left( - \frac{\left\|u-v\right\|^2}{2\sigma}\right) \\
%Two layer Neural Network & sigmoid \left(\eta\left(u\cdot v +1\right)+c\right) 
% \end{array} 
%\label{eq:diffkernel}
%\end{equation}

%The foundations of Support Vector Machines (SVM) have
\subsubsection{SVM  Classification}
The resulting SVM method can be summarized as follows 
\begin{enumerate}
	\item Selecting the SVM classifier parameters. Parameter $C$ representing the tradeoff between minimizing the training set error and maximizing the margin. 
	\item Select A kernel function and their kernel parameters. For example for the radial basis function kernel, one must select width of the Gaussian $\gamma$. 
	\item Solve dual QP Equation \ref{eq:dualkernel} or an alternative SVM formulation using an appropriate quadratic programming or linear programming algorithm. 
	\item Get the support vector and recover the imperial threshold  variable $b$. 
	\item Classify new points using $f(x)=sign\left(\sum{i}y_i\alpha_i K\left(x,x_i\right)-b\right)$
\end{enumerate}
The parameters in step 1. are selected using cross validation if sufficient data are available. %However, recent model  selectin strategies can 
\subsubsection {Multi Class Classifiers}

 There is mainly two popular methods to change the binary SVM classification problem into Multi class classification problem. The first methods is One Versus All (OVA) Strategy where M binary classifiers is constructed one classifier for each class.  The $ith$ classifier is trained using positive example form $ith$ class and negative example from all other classes. The output function $p_i$ represents the distance and classification of the example from the $ith$ classifier. For a new example $x$, OVA SVM strategy assigns it to the class with the largest value of $p_i$.

@@

In this step the user draw one or more stroke to draw a symbol. Figure \ref{fig:clock} illustrate the sequence and direction of user draw. In this section, each stroke user draws will be labeled as it illustrated in Figure \ref{fig:clockstroke1} and Figure \ref{fig:clockstroke2}.  Table \ref{tab:StepsStroke1} shows the detailed output of the program at each step. The input of this step is the a set of strokes, each stroke has a set of  points and the output is a list of segments that will be used in the recognition process. More detailed example and algorithms are demonstrated in the Appendix \ref{ChapterstepExample} \\

@@

\begin{enumerate}
\item Extracting $P_dp$ as in Section \ref{sec:Preprocessing} and Algorithm \ref{extractpdp}.
\item Ellipse Fitting: which is an attempt to fit as an ellipse as in details in Section \ref{sec:EllipseDetection} and Algorithm \ref{ellipseFitAlg} in Appendix \ref{ChapterstepExample}.
\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}. The algorithm will be similar to the PSO algorithm  in section \ref{sec:ParticleSwarmAlgorithm}. (see Algorithm \ref{AlgS1alg} in Appendix \ref{ChapterstepExample})
%he axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
\item Segmentation ALS2 as explained in Section \ref{sec:PolygonDivisionAlgorithm}. ( see Algorithm \ref{AlgS2alg} in Appendix \ref{ChapterstepExample} ). 
\item Append best segmentation to List of segmentation ($L_s$) of this symbols. 
\end{enumerate}

\begin{landscape}
The steps of the first stroke (details of the outputs are in  Appendix \ref{ChapterstepExample} in Table \ref{tab:StepsStroke1Details}):
\begin{scriptsize}
%\begin{table}
	%\centering 
%\scalebox{0.6}{		
\begin{longtable}{|p{2cm}|p{2cm}|p{5cm}|p{2cm}|p{5cm}|}
\caption{ Output of System in Each Step of Segmentation of First Stroke}
\label{tab:StepsStroke1} \\

@@

 \includegraphics[scale=0.5]{afterDefense/ellipse.jpg.eps}
 \\ \hline
 	\end{longtable}
 	\newpage
 	The steps of the second stroke:(details of the outputs are in  Appendix \ref{ChapterstepExample} in Table \ref{tab:StepsStroke1Details})

@@

	The input of the step is the list of segmentation $L_s$ that is used to compute the feature vector. Figure \ref{fig:segmentatiofinal.jpg} shows the best segmentation as given by the previous step. This step can be summarized as firstly using $L_s$ to computer the feature vector (Total of 70 features) and then introduce this vector the SVM classifier to classify it to one of the trained categories. The details of each step of feature extraction are detailed in the Algorithm \ref{FeatureAlg} . Feature extraction step is divided into four steps where each feature set is computed separately then all sets are appended to construct a final feature vector.

\begin{enumerate}
	\item Computing Structural and geometrical Features(FS1). This step computes number of parallel, perpendicular line and similar structural features [13 features]. Algorithm \ref{FeatureAlg} in Appendix \ref{ChapterstepExample}  is used to determine these features.
 	\item Rubine Feature Set (FS2) [12 features] .  Compute Rubine features as in \cite{gestureexample12} 
	\item Statistical Features (FS3) compute the moments as in \cite{zernike61} for N=10 there are 32 features.
	\item Global shape properties set (FS4). Firstly compute the convex hull of the stroke points then compute the rest of the features [13 features]. 
	\item Introduce the computed feature vector (All 70 features) to the classifier and check the label of the detected symbol. 
\end{enumerate}

@@

More details are in Table \ref{tab:StepsRegDetails} in Appendix \ref{ChapterstepExample}
